---
title: "Predicting the manner of exersize - PML course project"
output: html_document
---
Author: investh

18/01/2014

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

## Data processing

### Loading the libraries and the data
```{r, echo=TRUE}
    library(caret)
    library(kernlab)
    library(randomForest)

    train <- read.csv("pml-training.csv", na.strings=c("", "NA", "#DIV/0!"))
    test <- read.csv("pml-testing.csv", na.strings=c("", "NA", "#DIV/0!"))
```

### Building cross validation partitions
Splitting the data 70% for training, 30% for testing
```{r, echo=TRUE}
    inTrain <- createDataPartition(y=train$classe, p=0.7, list=FALSE)
    
    training <- train[inTrain,]
    testing <- train[-inTrain,]
    dim(training)
```

### Cleaning the data
Removing columns with a lot of NA(>70%),irrelevant columns and columns with near zero predictors
```{r, echo=TRUE}
    countNA <- apply(training,2,function(x) {sum(is.na(x))});
    training <- training[,which(countNA <  nrow(training)*0.7)];

    training <- training[,7:ncol(training)]

    nearZero <- nearZeroVar(train)
    train <- train[, -nearZero]
```

## Bulding the model and testing the accuracy

### Building random forest model
```{r, echo=TRUE}
    model <- randomForest(classe ~ ., data=training, ntrees=10)
```

### Validation the model accuracy on the testing set
```{r, echo=TRUE, message=TRUE}
    prediction <- predict(model, testing)
    print(confusionMatrix(prediction, testing$classe))
```
The results are great. The out of sample error is minimal.

## Predicting the values on the test set
```{r, echo=TRUE, message=TRUE}
testPrediction <- predict(model, test)
testPrediction
```

## writing the results to files for submission
```{r, echo=TRUE}
predictions <- as.vector(testPrediction)

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
            col.names = FALSE)
    }
}

pml_write_files(predictions)